### 2016-06-29 text/old/definitions.tex
Move old sections. Divorce paper from definitions file.

An expanded metamodel is an auxiliary construct where all the relations between types of a metamodel are made explicit, rather than remaining implicit in the specialization hierarchy. It is built by adding to the original metamodel typed graph a relation of type $t$ between two classes of the metamodel, whenever those classes specialize two classes that are also related by a relation of type $t$. Abstract classes and their relations do not carry over to the expanded metamodel. Expanded metamodels will be used in the subsequent text to facilitate formal the treatment of any structure involving polymorphism.

@@

A metamodel instance is a useful intermediate formal notion that lies between metamodel and model. The injective typed graph homomorphism between a metamodel instance and metamodel models multiple ``instances'' of objects and relations being typed by one single class or relation of the metamodel. Metamodel instances do not allow cyclic containment relations, as enforced by EMF.

@@

Given a metamodel instance, its containment transitive closure includes, besides the original graph, all the edges belonging to the transitive closure of containment links in that metamodel instance. The transitive edges are typed as \emph{indirect}. In the definitions that follow we will use the $*$ notation, as in \cref{def:instance_closure}, to denote the containment transitive closure of structures that directly or indirectly include metamodel instances. For example, $tg^{*}$ would represent the containment transitive closure of typed graph $tg$ wherever containment edges are found in the graph. Note that the $*$ notation is different from the $\star$ notation, introduced in \cref{def:expanded_metamodel} for an expanded metamodel.

@@

The expansion of a transformation rule is a set of transformation rules. Each rule in that set includes a possible replacement of each of the classes in the match part of the original rule by one of its subtypes. Expanded transformation rules will be important such that polymorphism is correctly handled in the developments that follow.

@@

%<*layertransformationtext>
\reviewer{I do not fully understand why this
  could be problematic. Do you have an example?}
\cref{def:layer_transformation} formalises the abstract syntax of a model transformation, introduced in \cref{sec:dsltrans}. An example of a model transformation can be observed in \cref{fig:dsltransformation}, the Police Station transformation. As expected, a formal DSLTrans transformation is composed of a sequence of layers where each layer is composed of a set of rules. The last condition of \cref{def:layer_transformation} imposes that, for any two pair of rules in the transformation, the matcher of the second rule never partially or totally subsumes (or contains) the matcher of the first rule, unless the second rule is in a subsequent layer and produces something more than the first rule. This condition avoids situations where the execution of a rule in a DSLTrans model transformation necessarily implies the execution of another rule (except for when rules having backward links necessarily execute because all their dependencies were created during the execution of previous layers).

@@

Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTrans transformation. Let also $pc = \langle V,E,st,\tau,Match,Apply,Rules\rangle \in
\textsc{Pathcond}(tr)$ of a path condition of $tr$ and $ex = \langle V',E',st',\tau',Input,Output\rangle \in Exec(tr)$ be an execution of $tr$. We have that $ex$ is abstracted by $pc$, noted $ex\Vvdash pc$, if and only if the set of transformation rules of $tr$ combined in $pc$ and the set of transformation rules of $tr$ used to built $ex$ is the same, and:
\begin{multline}
\label{eq:abstr_input_output}
\big(\forall rl\in Rules\;.\;Match(rl) \vartriangleleft Input^*\big) \;\land\; Output \blacktriangleleft Apply
\end{multline}
\center{and}
\begin{multline}
\label{eq:abstr_trace}
\big(\forall trc\in Components(pc|_{trace})\;.\;trc\vartriangleleft ex\big)\;\land\;\\
\big(\forall e'\in E'\, \exists e\in E\;.\;\tau'(e')=trace \Longrightarrow \tau(e)=trace\big)
\end{multline} 
% where for all $rl,rl'\in Rules$, if $rl\cong rl'$ then there exist two injective typed graph homomorphisms $f$ and $g$ such that $Match(rl)\stackrel{f}{\vartriangleleft} Input^*$, $Match(rl)\stackrel{g}{\vartriangleleft} Input^*$ and $CoDom(f)\neq CoDom(g)$.
% $/pc/ = Apply_{ex}\cup ex|_{trace}$ is the output part of $pc$, including symbolic traceability links.

@@

Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTrans transformation. Let also $pc = \langle V,E,st,\tau,Match,Apply,Rules\rangle \in
\textsc{Pathcond}(tr)$ of be a path condition of $tr$ and $ex = \langle V',E',st',\tau',Input,Output\rangle \in Exec(tr)$ be an execution of $tr$. We have that $ex$ is abstracted by $pc$, noted $ex\Vvdash pc$, if and only if the set of transformation rules of $tr$ combined in $pc$ and the set of transformation rules of $tr$ used to built $ex$ is the same, and:
\begin{multline}
\label{eq:abstr_input_output_appendix}
\big(\forall rl\in Rules\;.\;Match(rl) \vartriangleleft Input^*\big) \;\land\; Output \blacktriangleleft Apply
\end{multline}
\center{and}
\begin{multline}
\label{eq:abstr_trace_appendix}
\big(\forall trc\in Components(pc|_{trace})\;.\;trc\vartriangleleft ex\big)\;\land\;\\
\big(\forall e'\in E'\, \exists e\in E\;.\;\tau'(e')=trace \Longrightarrow \tau(e)=trace\big)
\end{multline} 
% where for all $rl,rl'\in Rules$, if $rl\cong rl'$ then there exist two injective typed graph homomorphisms $f$ and $g$ such that $Match(rl)\stackrel{f}{\vartriangleleft} Input^*$, $Match(rl)\stackrel{g}{\vartriangleleft} Input^*$ and $CoDom(f)\neq CoDom(g)$.
% $/pc/ = Apply_{ex}\cup ex|_{trace}$ is the output part of $pc$, including symbolic traceability links.

>>>>

An expanded metamodel is an auxiliary construct where all the relations between types of a metamodel are made explicit, rather than remaining implicit in the specialization hierarchy. It is built by adding to the original metamodel typed graph a relation of type $t$ between two classes of the metamodel, whenever those classes specialize two classes that are also related by a relation of type $t$. Abstract classes and their relations do not carry over to the expanded metamodel. Expanded metamodels will be used in the subsequent text to facilitate formal treatment of any structure involving polymorphism.

@@

A metamodel instance is a useful intermediate formal notion that lies between metamodel and model. The injective typed graph homomorphism between a metamodel instance and metamodel models multiple ``instances'' of objects and relations being typed by one single class or relation of the metamodel. Metamodel instances do not allow cyclic containment relations, as enforced by EMF. \reviewer{Do you mean “in analogy to EMF”?}

@@

Given a metamodel instance, its containment transitive closure includes, besides the original graph, all the edges belonging to the transitive closure of containment links in that metamodel instance. The transitive edges are typed as \emph{indirect}. In the definitions that follow we will use the $*$ notation, as in \cref{def:instance_closure}, to denote the containment transitive closure of structures that directly or indirectly include metamodel instances. For example, $tg^{*}$ would represent the containment transitive closure of typed graph $tg$ wherever containment edges are found in the graph. Note that the $*$ notation is different from the $\star$ notation, introduced in \cref{def:expanded_metamodel} for an expanded metamodel. \reviewer{The difference between these stars is too small to be noticeable.}

@@

The expansion of a transformation rule is a set of transformation rules. Each rule in that set includes a possible replacement of each of the classes in the match part of the original rule by one of its subtypes. Expanded transformation rules will be important to ensure that polymorphism is correctly handled in the developments that follow.

@@

%<*layertransformationtext>

\cref{def:layer_transformation} formalises the abstract syntax of a model transformation, introduced in \cref{sec:dsltrans}. An example of a model transformation can be observed in \cref{fig:dsltransformation}, the Police Station transformation. As expected, a formal DSLTrans transformation is composed of a sequence of layers where each layer is composed of a set of rules. The last condition of \cref{def:layer_transformation} imposes that, for any two pair of rules in the transformation, the matcher of the second rule never partially or totally subsumes (or contains) the matcher of the first rule, unless the second rule is in a subsequent layer and produces something more than the first rule. This condition avoids situations where the execution of a rule in a DSLTrans model transformation necessarily implies the execution of another rule (except for when rules having backward links necessarily execute because all their dependencies were created during the execution of previous layers). \reviewer{I do not fully understand why this
  could be problematic. Do you have an example?}

@@

Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTrans transformation. Let also $pc = \langle V,E,st,\tau,Match,Apply,Rules\rangle \in
\textsc{Pathcond}(tr)$ of a path condition of $tr$ and $ex = \langle V',E',st',\tau',Input,Output\rangle \in Exec(tr)$ be an execution of $tr$. We have that $ex$ is abstracted by $pc$, noted $ex\Vvdash pc$, if and only if the set of transformation rules of $tr$ combined in $pc$ and the set of transformation rules of $tr$ used to build $ex$ is the same, and:
\begin{multline}
\label{eq:abstr_input_output}
\big(\forall rl\in Rules\;.\;Match(rl) \vartriangleleft Input^*\big) \;\land\; Output \blacktriangleleft Apply
\end{multline}
\center{and}
\begin{multline}
\label{eq:abstr_trace}
\big(\forall trc\in Components(pc|_{trace})\;.\;trc\vartriangleleft ex\big)\;\land\;\\
\big(\forall e'\in E'\, \exists e\in E\;.\;\tau'(e')=trace \Longrightarrow \tau(e)=trace\big)
\end{multline} 
% where for all $rl,rl'\in Rules$, if $rl\cong rl'$ then there exist two injective typed graph homomorphisms $f$ and $g$ such that $Match(rl)\stackrel{f}{\vartriangleleft} Input^*$, $Match(rl)\stackrel{g}{\vartriangleleft} Input^*$ and $CoDom(f)\neq CoDom(g)$.
% $/pc/ = Apply_{ex}\cup ex|_{trace}$ is the output part of $pc$, including symbolic traceability links.

@@

Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTrans transformation. Let also $pc = \langle V,E,st,\tau,Match,Apply,Rules\rangle \in
\textsc{Pathcond}(tr)$ of be a path condition of $tr$ and $ex = \langle V',E',st',\tau',Input,Output\rangle \in Exec(tr)$ be an execution of $tr$. We have that $ex$ is abstracted by $pc$, noted $ex\Vvdash pc$, if and only if the set of transformation rules of $tr$ combined in $pc$ and the set of transformation rules of $tr$ used to build $ex$ is the same, and:
\begin{multline}
\label{eq:abstr_input_output_appendix}
\big(\forall rl\in Rules\;.\;Match(rl) \vartriangleleft Input^*\big) \;\land\; Output \blacktriangleleft Apply
\end{multline}
\center{and}
\begin{multline}
\label{eq:abstr_trace_appendix}
\big(\forall trc\in Components(pc|_{trace})\;.\;trc\vartriangleleft ex\big)\;\land\;\\
\big(\forall e'\in E'\, \exists e\in E\;.\;\tau'(e')=trace \Longrightarrow \tau(e)=trace\big)
\end{multline} 
% where for all $rl,rl'\in Rules$, if $rl\cong rl'$ then there exist two injective typed graph homomorphisms $f$ and $g$ such that $Match(rl)\stackrel{f}{\vartriangleleft} Input^*$, $Match(rl)\stackrel{g}{\vartriangleleft} Input^*$ and $CoDom(f)\neq CoDom(g)$.
% $/pc/ = Apply_{ex}\cup ex|_{trace}$ is the output part of $pc$, including symbolic traceability links.

### 2016-06-29 text/property_verification.tex
Move old sections. Divorce paper from definitions file.

\begin{definition}{Property of a Transformation\\}
\label{def:trans_prop}
\CatchFileBetweenTags{\propertydef}{text/definitions}{propertydef}{\propertydef}
\end{definition}

@@

\begin{definition}{Satisfaction of a Property by an Execution of a Transformation\\}
\label{def:sat_prop_ex}
\CatchFileBetweenTags{\satpropexe}{text/definitions}{satpropexe}{\satpropexe}
\end{definition}

@@

\begin{definition}{Satisfaction of a Property by a Path Condition\\}
\label{def:sat_prop_pc}
\CatchFileBetweenTags{\satproppc}{text/definitions}{satproppc}{\satproppc}
\end{definition}

>>>>

\begin{definition}{Property of a Transformation\\}
\label{def:trans_prop}

\reviewer{You can not handle properties with multiple instances of rule 
    element (example in Fig. 25). On p.14 in the answer section,
    you claim that such conditions are excluded in Def. 29. However,
    the last lines of this definition are not easy to understand.
    Please explain why there are there (by giving the explanation
    from the answers section.) Without these explanations it is very hard to see why 
    Prop. 3 should hold.}

Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTrans transformation. A property of $tr$ is a 6-tuple $\langle V,E,(s,t),\tau,Pre,Post\rangle$, where $Pre = \langle V',E',st',\tau'\rangle \in \textsc{Ipattern}^{sr}$ and $Post = \langle V'',E'',st''\tau''\rangle \in\\ \textsc{Ipattern}^{tg}$ are indirect metamodel patterns. We also have that $V=V'\cup V''$, $E\subseteq E'\cup E''$ and $\tau\subseteq \tau'\cup \tau''$, where the co-domain of $\tau$ is the union of the co-domains of $\tau'$ and $\tau''$ and the set $\{trace\}$. An edge $e \in E\setminus E'\cup E''$ is called a \emph{traceability link} and is such that $s(e)\in V''$, $t(e)\in V'$ and $\tau(e)=trace$. Finally we have that there is at least one path condition $\big\langle V_{pc},E_{pc},st_{pc},\tau_{pc}, Match,Apply,Rule\big\rangle\in \textsc{Pathcond}(tr)$ for which a surjective typed graph homomorphism $m\stackrel{f}{\blacktriangleleft} Pre$ exists, where $m\sqsubseteq Match$ and $f(v)\neq f(v')$ if $v$ and $v'$ are elements of the path condition belonging to the same rule of set $Rule$. The set of all properties of transformation $tr$ is called $\textsc{Property}(tr)$.

\end{definition}

@@

\begin{definition}{Satisfaction of a Property by an Execution of a Transformation\\}
\label{def:sat_prop_ex}

Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a transformation. Let also $p = \langle V,E,st,\tau,\\Pre,Post\rangle \in \textsc{Property}(tr)$ be a property of $tr$ and $ex = \langle V',E',\\st',\tau',Input,Output\rangle \in Exec(tr)$ be an execution of $tr$. Execution $ex$ satisfies property $p$,  written $ex \models p$, if and only if:
\begin{multline*}
\forall f\; \exists g\,.\,\big(Pre \stackrel{f}{\vartriangleleft} Input^{*} \implies p \stackrel{g}{\vartriangleleft} ex^*\big)\\\text{where } V(Input)\cap CoDom(g) = CoDom(f)
\end{multline*}
% \begin{enumerate}
% \item $\langle V_p,E_p\setminus Il_p,\tau_p\rangle$ is a typed subgraph of $\langle V_x, E_x,\tau_x\rangle$
% \item if $v_p\rightarrow v_{p}'\in Il_p$ then there exists $v_x\rightarrow
% v_{x}'\in E_{x}^{*}$ where $\tau(v_p)=\tau(v_x)$, $\tau(v_p')=\tau(v_x')$ and
% $E_{x}^{*}$ is obtained by the transitive closure of $E_{x}$.
% \end{enumerate}
\end{definition}

@@

\begin{definition}{Satisfaction of a Property by a Path Condition\\}
\label{def:sat_prop_pc}

\reviewer{In Def. 31 I found it confusing that - different from previous
    definitions - the two morphisms now go in the same direction.
    There is an explanation after Definition 31, but I found it 
    very hard to understand.}
Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a transformation. Let also $p = \langle V,E,st,\tau,\\Pre,Post\rangle \in \textsc{Property}(tr)$ be a property of $tr$ and $pc=\big\langle V',E',\\st',\tau', Match,Apply,Rulecop\big\rangle\in \textsc{Pathcond}(tr)$ be a path condition of $tr$. Path condition $pc$ satisfies property $p$, written $pc \vdash p$, if and only if:
\begin{multline*}
\forall f\; \exists g\,.\,\big(in\stackrel{f}{\blacktriangleleft} Pre \implies out \stackrel{g}\blacktriangleleft p\big)\\
\text{where } in \sqsubseteq Match^{*} \;\land\; out \sqsubseteq pc^{*}
\end{multline*}
Additionally $Dom(g)\cap Match(pc^{*}) = Dom(f)$ and $f(v)\neq f(v')$, $g(v)\neq g(v')$ whenever $v$ and $v'$ are elements of the path condition belonging to the same rule copy of set $Rulecop$.

\end{definition}
